---
title: "Project 3: Tutorial"
author: "Alexander Knowlton"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This is a package that holds several useful functions for statistical
inference and prediction, including a one-sample t test, a linear regression,
a random forest function for penguins data, and a cross-validation function.

To install and load the package from GitHub, use the following code:
```{r eval = FALSE}
devtools::install_github("Codax2000/mypackage")
library(mypackage)
```

```{r echo = FALSE}
library(mypackage)
```

## my_t.test
`my_t.test` is a function that performs a one-sample t test on a set of given
data. It accepts alternative hypotheses and a value for mu, the null parameter.
It returns a `list` with values of the p value, the alternative hypothesis,
the t statistic, and the degrees of freedom of the data.

For example, to test the hypotheses $$H_0: \mu=60$$ $$H_a: \mu \neq 60$$ on the
`lifeExp` column of the `my_gapminder` data, use the following code:
```{r}
local_gapminder = na.omit(my_gapminder)
result <- my_t.test(local_gapminder$lifeExp, alternative = "two.sided", mu = 60)
print(result)
```
Since the p-value of the data is `r round(result$p_val, 3)`, this test indicates
that, at $\alpha=0.05$, there is no evidence that the true mean
life expectancy of the gapminder data is not equal to 60.

To test the hypotheses $$H_0: \mu=60$$ $$H_a: \mu \lt 60$$ on the
`lifeExp` column of the `my_gapminder` data, use the following code:
```{r}
local_gapminder = na.omit(my_gapminder)
result <- my_t.test(local_gapminder$lifeExp, alternative = "less", mu = 60)
print(result)
```
Since the p-value of the data is `r round(result$p_val, 3)`, this test indicates
that, at $\alpha=0.05$, there is evidence that the true mean
life expectancy of the gapminder data is less than 60.

And finally, to test the hypotheses that $$H_0: \mu=60$$ $$H_a: \mu \gt 60$$
```{r}
local_gapminder = na.omit(my_gapminder)
result <- my_t.test(local_gapminder$lifeExp, alternative = "greater", mu = 60)
print(result)
```
Since the p-value of the data is `r round(result$p_val, 3)`, this test indicates
that, at $\alpha=0.05$, there is no evidence that the true mean
life expectancy of the gapminder data is greater than 60.

## my_lm
`my_lm` is a function that fits a linear model to a data set. It takes a data
set and a formula as a parameter and returns a
matrix with the coefficients of each predictor, the t statistic for each one,
the standard error of each coefficient, and the p value of each one, which
tells the significance of that coefficient.

For example, we
can fit a model to the `gapminder` data to try and predict life expectancy from
gdp per capita and the continent.

To fit the model, use the following:
``` {r}
local_gapminder <- na.omit(my_gapminder)
result <- my_lm(local_gapminder, lifeExp ~ gdpPercap + continent)
print(result)
```

the coefficient for `gdpPercap` is `r round(result[2, 1], 4)`. This indicates
that, if `continent` is kept constant, an increase of 1 in `gdpPercap` will
mean an increase of `r round(result[2, 1], 4)` years in life
expectancy.
